### 2019-12-12 Draft/Table1.tex

\begin{table}[H]
\begin{center}
\bgroup
\def\arraystretch{1.2}
\begin{tabular}{|l|l|l|}
\hline
\hline
\textbf{Quantity} & \textbf{\tit{short}$-\Tcoh$ \tit{setup}} & \textbf{\tit{long}$-\Tcoh$ \tit{setup}}\\
\hline
\hline
$\mathrm{T}_{\mathsf{coh}}$ (hours) & 60 & 210 \\ \hline
$\mathrm{N}_{\mathsf{seg}}$ & 64 & 12       \\ \hline
$\mathrm{T}_{\mathsf{obs}}$ (hours) & $60\times64$ & $210\times12$       \\ \hline
$\delta\!f$ (Hz) & $6.7\times 10^{-7}$ & $3.3\times 10^{-7}$\\ \hline
$\delta\!\dot{f}$ (Hz/s) & $1.3\times 10^{-13}$ & $1.6\times 10^{-13}$ \\  \hline
$\mathrm{d}_\mathsf{sky}$ &  0.024 & 0.068\\
%$\mathrm{t}_\mathsf{ref}$ (\textsf{GPS} sec) & 1132729647.5 & 1170946472.0 \\
\hline
\hline
\end{tabular}
\egroup
\end{center}
\caption{The short$-\Tcoh$ and long$-\Tcoh$ search setups.}
\label{table:setup}
\end{table}